%!TEX root = ../thesis.tex

%This is the Summary
%%=========================================
\cleardoublepage
\addcontentsline{toc}{section}{Acknowledgments}
\section*{Acknowledgments}
In the name of Allah, the Most Gracious, the Most Merciful.

With the completion of this dissertation, I am grateful to have fulfilled a long-standing dream of mine.
From an early age, I have aspired to follow in the footsteps of esteemed scholars such as Al-Khwarizmi, Ibn Sina, and Qadi Zada al-Rumi.
I would like to express my sincerest gratitude to Prof. Dr. David Bermbach, who has been a tremendous support in helping me achieve this goal.
I am deeply grateful to have had the opportunity to work, research, learn, and teach alongside him.
I am grateful to have had the opportunity to learn from him not only how to become a better scientist, but also in many other areas of life.
He went above and beyond what I could have asked for.
I believe every scientist would benefit from having such a great mentor.

In addition to Prof. Dr. Bermbach, I would also like to express my gratitude to the chair of the doctoral committee, Prof. Dr. Ziawasch Abedjan, as well as the other reviewers of the doctoral committee, Prof. Dr. Heather Kaths and Prof. Dr. Odej Kao.
I am grateful to have received constructive feedback from them, which has instilled confidence in my own research and demonstrated ways to present significant contributions in a more effective and consistent manner.

I would also like to express my gratitude to my colleagues in the Scalable Software Systems research group, my esteemed co-authors, and my honored students, to whom I was privileged to be able to offer guidance as a supervisor.
I am grateful to have had your support and encouragement, which have been invaluable in helping me achieve success in my work.

I would also like to express my gratitude to my family. My parents have always believed in me, and I am thankful to have them as a source of strength.
My wife, sister, and daughter have also been a great source of support, even when I was away pursuing my studies.

%I am also grateful to the many people whose advice and education I have enjoyed, who have played a part in this result. 
%I am grateful to my relatives, educators, and teachers.
%I consider myself fortunate to have had the opportunity to learn from and be mentored by many excellent individuals.

All praise is for Allah -- Lord of everything in existence.  

\afterpage{\null\thispagestyle{empty}\newpage}
